Tidy data structure to support exploration and modeling of temporal-context data

### 2018-11-17 main.Rmd
added transformation verbs paragraph

What is notable about data pipeline is that it not only frees users from a tedious and error-prone analysis but also empowers a wider audience to focus on the data analysis hand without concerning the details of computational implementation. A fluent and fluid pipeline glues tidy data and the grammar of data manipulation together. It helps (1) break up a big problem to into manageable blocks, (2) generate human readable analysis workflow, (3) avoid introducing mistakes as many as possible.

@@

This tidy data representation choice will prove suitable for almost all conceivable applications of temporal data. Time-based pipeline starts with raw data however it lands. It is advised to scrutinize identical entries of key and index in the right beginning. Duplicates signal the data quality issue, which would likely affect the following analyses and hence decision making. Analysts are encouraged to gaze at data as earlier as possible and reason about the process of data cleaning. When the data meets the tsibble standard, it flows into the exploration stage.

@@

## Clarity and transparency

Transformation assembles a set of modules. A unit that makes up a module is function or preferably "verb". A tsibble is an object, considered as a conceptualized and materialized noun, an action performed on the object can be phrased as a verb. Naming verbs is difficult. This verb is self-explanatory enough to express what it is supposed to do or fail. We shall adapt **dplyr** key verbs to time domain and expand its vocabulary a little more to handle time-related analysis problems.

* **implicit missingness**: `has_gaps()`, `count_gaps()`, `fill_na()`
* **row-wise**: `filter()`, `slice()`, `arrange()`
* **column-wise**: `mutate()`, `select()`, `summarise()`
* **group-wise**: `index_by()`, `group_by()`
* **reshape**: `gather()`, `spread()`, `nest()`, `unnest()`
* **rolling window**: `slide()`, `tile()`, `stretch()`

## Composition and modularity

@@

>>>>

What is notable about data pipeline is that it not only frees users from a tedious and error-prone analysis but also empowers a wider audience to focus on the data analysis without concerning the details of computational implementation. A fluent and fluid pipeline glues tidy data and the grammar of data manipulation together. It helps (1) break up a big problem to into manageable blocks, (2) generate human readable analysis workflow, (3) avoid introducing mistakes as many as possible.

@@

This tidy data representation choice will prove suitable for almost all conceivable applications of temporal data. Time-based pipeline starts with raw data however it arrives. It is advised to scrutinize identical entries of key and index in the right beginning. Duplicates signal the data quality issue, which would likely affect the following analyses and hence decision making. Analysts are encouraged to gaze at data as earlier as possible and reason about the process of data cleaning. When the data meets the tsibble standard, it flows into the exploration stage.

@@

Transformation assembles a set of modules. A unit that makes up a module is function or preferably "verb". A tsibble is an object, conceptually considered as a noun, and hence an action performed on the object can be phrased as a verb. Each verb focuses on one thing and achieve its goal. The verb should be self-explanatory to advise what it is supposed to do or fail, for example `filter()` picks observations, and `select()` picks variables. These general-purpose verbs are made available in the **tidyvere** suite. When manipulating in temporal context, these verbs are adapted to time domain. A perceivable difference is summarizing variables between data frame and tsibble. The former will reduce to a single summary, whereas the latter will retain the index and their corresponding summaries. New tsibble-specific verbs are proposed to expand the vocabulary in order to facilitate temporal data analysis.

A table to list the verbs and what they do.

* **implicit missingness**: `has_gaps()`, `count_gaps()`, `fill_na()`
* **row-wise**: `filter()`, `slice()`, `arrange()`
* **column-wise**: `mutate()`, `select()`, `summarize()`
* **group-wise**: `index_by()`, `group_by()`
* **reshape**: `gather()`, `spread()`, `nest()`, `unnest()`

* Clarity and transparency
* Composition and modularity

@@

## Rolling window in functional programming

* **rolling window**: `slide()`, `tile()`, `stretch()`
